### 2013-03-01 doc/Sigmoid_Functions.tex

\documentclass{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{color}
\usepackage{subfig}
\usepackage{float} % For [H] placement

\usepackage[ruled,vlined,linesnumbered]{algorithm2e}

\usepackage{booktabs}

\usepackage[bookmarks=true,ocgcolorlinks=true,plainpages=false, breaklinks=true, bookmarksopen=true, bookmarksnumbered=true]{hyperref}
%\hypersetup{bookmarks=false}  %hide the bookmarks bar
%\hypersetup{bookmarksopen=false}  % expand tree of bookmarks or just show first level
\hypersetup{linkcolor=blue, citecolor=magenta,urlcolor=blue} % electronic
\hypersetup{colorlinks=true}

\usepackage[T1]{fontenc}
\usepackage{palatino}

\allowdisplaybreaks

\renewcommand{\vec}[1]{\ensuremath{{\boldsymbol #1}}}
\newcommand{\mat}[1]{\ensuremath{\boldsymbol{#1}}}

\DeclareMathOperator*{\erf}{erf}

\title{PCRB}

\date{\today}

\begin{document}

\section{Sigmoid functions}

The sigmoid can be specified using the logistic function
\begin{align}
	g(x) = \frac{1}{1 + \exp(\varsigma(v_0 - x))}
\end{align}
which has the derivative
\begin{align}
	g'(x) = \varsigma g(x)(1 - g(x))
\end{align}

Alternatively, using the error function
\begin{align}
	g(x) = \frac{1}{2}\left( 1 + \erf\left( \frac{x - v_0}{\xi\sqrt{2}}\right)\right)
\end{align}
which has the derivative
\begin{align}
	g'(x) = \frac{1}{\xi\sqrt{2\pi}} \exp\left(-\left(\frac{x - v_0}{\xi\sqrt{2}}\right)^2 \right) 
\end{align}

\section{Relationship between widths}

Through numerical simulation we can deduce that setting the error function width to
\begin{align}
	\xi = \frac{1.699}{\varsigma}
\end{align}
gives a sigmoid with approximately the same shape as a logistic function with width $\varsigma$. This relationship seems to hold across a wide range of widths. It is likely this number could be derived analytically.

% \bibliography{../../library}

\end{document}